---
title: "Homework 1"
format:
  html:
    self-contained: true
    code-fold: true
editor: visual
---
This is the link to the repo : <https://github.com/wei-min-lin/Homework1.git>.

# Q1

## a

```{r}
UCI <- read.csv('/Users/weimin/Documents/Class Materials/STATS 506 R/forty+soybean+cultivars+from+subsequent+harvests/data.csv')
is.data.frame(UCI)
```

## b

```{r}
summary(UCI)
length(table(UCI$Cultivar))
table(UCI$Season)
#summary(UCI) show that the total observation number are 320.
#length(table(UCI$Cultivar)) shows that there are 40 distinct species.
#table(UCI$Season) show that there are only two different seasons.
```

##c

```{r}
# install.packages("dplyr")
library(dplyr)
# variable define
UCI <- UCI %>%
  rename(
    plant_height = PH,
    grain_yield = GY,
    season = Season
  )
# Calculate correlation

UCI %>%
   summarise(
    correlation = cor(
      plant_height,
      grain_yield
    )
  )


# Calculate correlation (by season)
UCI %>%
  group_by(season) %>%
  summarise(
    correlation = cor(
      plant_height,
      grain_yield
    )
  )


# 3. What is the grain yield of the soybean observation with the greatest plant height?


UCI$grain_yield[which.max(UCI$plant_height)]


# 4. What percentage of observations have a thousand seed weight (MHG) greater than the overall average thousand seed weight?

table(UCI$MHG > mean(UCI$MHG))
151 / (151+169)


# 5. Which cultivar has the highest average grain yield?

which.max(tapply(UCI$grain_yield, UCI$Cultivar, mean))
```

### (1.)

The correlation between plant height and grain yield is 0.1233

### (2.)

The higher correlation season is Season 1.

Season 1's correlation : 0.2309

Season 2's correlation : -0.0227

### (3.)

3304.678

### (4.)

47.19%

### (5.)

98R30 CE 
        
## d

```{r}
numbers <- UCI[sapply(UCI,is.numeric)]


overall <- c("Overall", colMeans(numbers, na.rm = TRUE))
Season <- aggregate(numbers, by = list(UCI$season), mean, na.rm = TRUE)

A = rbind(overall, Season)
A
```

## e

```{r}
t.test(grain_yield ~ season, data = UCI)
```

Since the p-value is greater than 0.05, we fail to reject the null hypothesis. Therefore, there is not sufficient evidence to conclude that the mean grain yield differs between the two seasons.

The 95% confidence interval for the difference in means is (-91.157, 130.752), which includes 0, further supporting the conclusion that there is no statistically significant difference between the two seasons.


# Q2

## a

```{r}
# 1. Import data
adult <- read.csv('/Users/weimin/Documents/Class Materials/STATS 506 R/adult/adult.data',
                  header = FALSE,
                  strip.white = TRUE,
                  na.strings = "?")
```

## b

```{r}
# 2. Variable names
names(adult) <- c(
  "age", "workclass", "fnlwgt", "education", "education_num",
  "marital_status", "occupation", "relationship", "race", "sex",
  "capital_gain", "capital_loss", "hours_per_week",
  "native_country", "income"
)
```


## c

```{r}
colSums(is.na(adult))
```
There are 1836 missing value in workclass, 1843 in occupation, and 583 in native_country

## d
```{r}
adult <- adult[adult$hours_per_week >= 10,]
```

It had 32561 before filtered, and it has 32103 after filtered.

## e
```{r}
adult <- adult[
  !is.na(adult$occupation) &
  !is.na(adult$workclass) &
  !is.na(adult$native_country), ]
head(adult)
```

Observations with missing occupation, workclass, or native country were removed because these observations do not contain complete information for these categorical variables.

## f
```{r}
adult <- adult[adult$age >= 18 & adult$age <= 90 & adult$education != "Preschool",]
```

## g
```{r}
nrow(adult)
```

The final sample size is 29523.

# Q3
## a

```{r}
nextHappy <- function(n) {

  if (!is.numeric(n) || length(n) != 1 || is.na(n) ||
      !is.finite(n) || n <= 0 || n %% 1 != 0) {
    stop("n must be a positive integer.")
  }

  digits <- as.numeric(strsplit(as.character(n), "")[[1]])
  sum(digits^2)
}

nextHappy(82)
```


## b

```{r}
happySequence <- function(n) {
  # Check input
  if (!is.numeric(n) || length(n) != 1 || is.na(n) ||
      !is.finite(n) || n <= 0 || n %% 1 != 0) {
    stop("n must be a positive integer.")
  }

  sequence <- n
  seen <- numeric(0)

  while (TRUE) {
    current <- tail(sequence, 1)

    # Happy number
    if (current == 1) {
      return(list(
        sequence = sequence,
        isHappy = TRUE
      ))
    }

    # Cycle detected
    if (current %in% seen) {
      return(list(
        sequence = sequence,
        isHappy = FALSE
      ))
    }

    seen <- c(seen, current)
    sequence <- c(sequence, nextHappy(current))
  }
}

happySequence(4)

happySequence(19)
```
## c

### a

```{r}
results <- lapply(1:1000, happySequence)

isHappy <- sapply(results, function(x) x$isHappy)

numHappy <- sum(isHappy)
numHappy
```

143

### b

```{r}
percentHappy <- mean(isHappy) * 100
percentHappy
```

14.3%


### c
```{r}
sequenceLength <- sapply(
  results,
  function(x) {
    if (x$isHappy) {
      length(x$sequence)
    } else {
      NA
    }
  }
)

maxLength <- max(sequenceLength, na.rm = TRUE)
maxLength

longest <- which(sequenceLength == maxLength)
longest
```

The longest happy sequence is 7, and it produced by 21 starting values.

356 365 379 397 469 496 536 563 635 649 653 694 739 793 899 937 946 964 973 989 998


