---
title: "STATS 506 Homework2"
author: "Wei min"
format:
  pdf:
    pdf-engine: xelatex
    documentclass: article
    fontsize: 10pt
    geometry: margin=0.8in
    colorlinks: true
  html:
    embed-resources: true
    code-fold: false
    toc: true
number-sections: false
execute:
  echo: true
  warning: true
  message: false
editor: source
---

```{r}
#| include: false

knitr::opts_chunk$set(fig.width = 7, fig.height = 4.5, comment = NA)
library(microbenchmark)
```

# Problem 1: Random Walk with Momentum

## (a)

I keep track of the current direction and add each step to the position.
When a new direction is drawn, it can be the same as the previous one.

```{r}
momentum_walk <- function(n, p) {
  stopifnot(length(n) == 1, is.numeric(n), is.finite(n),
            n >= 1, n == floor(n),
            length(p) == 1, is.numeric(p), is.finite(p),
            p >= 0, p <= 1)

  steps <- rbind(c(1, 0), c(-1, 0), c(0, 1), c(0, -1))
  direction <- sample(1:4, 1)
  position <- c(0, 0)

  for (i in seq_len(n)) {
    if (i > 1 && runif(1) >= p) {
      direction <- sample(1:4, 1)
    }
    position <- position + steps[direction, ]
  }
  position
}

set.seed(1101)
momentum_walk(10, 0.6)
```

`stopifnot()` checks that `n` is a positive integer and `p` is between 0 and 1.
The output is the final position, `(x, y)`.

## (b)

The function above uses a loop, so I use it as `momentum_walk1`.

```{r}
momentum_walk1 <- momentum_walk

momentum_walk2 <- function(n, p) {
  stopifnot(length(n) == 1, is.numeric(n), is.finite(n),
            n >= 1, n == floor(n),
            length(p) == 1, is.numeric(p), is.finite(p),
            p >= 0, p <= 1)

  steps <- rbind(c(1, 0), c(-1, 0), c(0, 1), c(0, -1))
  direction <- sample(1:4, 1)

  moves <- sapply(seq_len(n), function(i) {
    if (i > 1 && runif(1) >= p) {
      direction <<- sample(1:4, 1)
    }
    steps[direction, ]
  })

  rowSums(moves)
}

set.seed(1102)
momentum_walk1(10, 0.6)
momentum_walk2(10, 0.6)
momentum_walk1(1000, 0.6)
momentum_walk2(1000, 0.6)
```

The loop adds each step directly to the position. The second version uses
`sapply()` to store the steps, then adds them with `rowSums()`.
Here, `<<-` updates `direction` inside `momentum_walk2` so it carries over
to the next step.

## (c)

```{r}
set.seed(1103)
momentum_walk1(10, 0.6)
set.seed(1103)
momentum_walk1(10, 0.6)

set.seed(1104)
momentum_walk1(1000, 0.6)
set.seed(1104)
momentum_walk1(1000, 0.6)
```

Each pair gives the same result. Resetting the seed makes R use the same
sequence of random numbers again.

## (d)

```{r}
set.seed(1105)
p_values <- c(0, 0.5, 0.9)

distances <- sapply(p_values, function(p) {
  position <- momentum_walk1(1000, p)
  sqrt(sum(position^2))
})

data.frame(p = p_values, distance = distances)
```

The distances are about 50.54, 56.30, and 243.82. They increase with momentum
in these runs, but one walk at each value of `p` is not enough to tell whether
this usually happens.

## (e)

```{r}
set.seed(1106)
p_values <- c(0, 0.25, 0.5, 0.75, 0.9)

mean_distance <- sapply(p_values, function(p) {
  distances <- replicate(1000, {
    position <- momentum_walk1(1000, p)
    sqrt(sum(position^2))
  })
  mean(distances)
})

data.frame(p = p_values, mean_distance = mean_distance)
```

The mean distance increases from about 27.9 at `p = 0` to 124.3 at `p = 0.9`.
With higher momentum, the walk stays in the same direction longer, so it
usually ends farther from the origin.

## (f)

```{r}
#| warning: false

set.seed(1107)
small <- microbenchmark(
  loop = momentum_walk1(1000, 0.6),
  apply = momentum_walk2(1000, 0.6),
  times = 30
)
large <- microbenchmark(
  loop = momentum_walk1(100000, 0.6),
  apply = momentum_walk2(100000, 0.6),
  times = 30
)

timing <- rbind(
  data.frame(n = 1000, summary(small, unit = "ms")[, c("expr", "median")]),
  data.frame(n = 100000, summary(large, unit = "ms")[, c("expr", "median")])
)
names(timing) <- c("n", "method", "median_ms")
knitr::kable(timing, digits = 3, row.names = FALSE)
```

The loop is faster at both sizes in this run, and the time gap is larger at
100,000 steps. `sapply()` calls a function for every step and stores all the
moves before adding them. The loop only updates the current position, so it
does less extra work. The exact timings will vary between runs.

\newpage

# Problem 2: Campus Bus Arrivals

```{r}
set.seed(2201)
days <- 10000

morning <- rowSums(matrix(rpois(days * 3, 4), nrow = days))
peak_9am <- rnorm(days, mean = 45, sd = 4)
midday <- rowSums(matrix(rpois(days * 6, 15), nrow = days))
peak_4pm <- rnorm(days, mean = 55, sd = 5)
evening <- rowSums(matrix(rpois(days * 6, 9), nrow = days))

total <- morning + peak_9am + midday + peak_4pm + evening
mean(total)
```

The estimated average is about **255.88 passengers per day**. As a check,
the expected total is

$$
3(4)+45+6(15)+55+6(9)=256.
$$

I used standard deviations 4 and 5 for the Normal draws, since the given
variances are 16 and 25. The code simulates independent hourly counts for
all days at once, without loops.

\newpage

# Problem 3: Linear Regression

## (a)

```{r}
url <- paste0(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/",
  "master/data/2021/2021-03-02/youtube.csv"
)
youtube <- read.csv(if (file.exists("youtube.csv")) "youtube.csv" else url)

remove <- c("brand", "superbowl_ads_dot_com_url", "youtube_url",
            "id", "etag", "published_at", "title", "description",
            "thumbnail", "channel_title")
youtube <- youtube[, !names(youtube) %in% remove]
dim(youtube)
```

The cleaned data have **247 rows and 15 columns**. I removed the names,
links, IDs, publication timestamp, and other fields that could identify an
ad. I kept `year` for the regressions, along with `kind` and `category_id`,
which are general categories rather than identifiers.

## (b)

```{r}
engagement <- c("view_count", "like_count", "dislike_count",
                "favorite_count", "comment_count")
summaries <- data.frame(
  outcome = engagement,
  observed = sapply(youtube[engagement], function(x) sum(!is.na(x))),
  zeros = sapply(youtube[engagement], function(x) sum(x == 0, na.rm = TRUE)),
  median = sapply(youtube[engagement], median, na.rm = TRUE),
  mean = sapply(youtube[engagement], mean, na.rm = TRUE),
  max = sapply(youtube[engagement], max, na.rm = TRUE)
)
knitr::kable(summaries, digits = 1, row.names = FALSE)
```
```{r}
#| fig-height: 6
par(mfrow = c(3, 2), mar = c(3.5, 3.5, 2, 1),
    mgp = c(1.8, 0.55, 0), cex.lab = 0.85)
for (variable in engagement) {
  x <- na.omit(youtube[[variable]])
  if (variable == "favorite_count") {
    barplot(table(x), main = variable, xlab = "Count",
            ylab = "Frequency", col = "lightblue")
  } else {
    hist(x, breaks = 25, main = variable, xlab = "Count",
         col = "lightblue", cex.axis = 0.75)
  }
}
par(mfrow = c(1, 1))
```
| Outcome | Choice | Reason |
|:--|:--|:--|
| Views | Use `log(1 + count)` | A few ads have far more views than the rest. |
| Likes | Use `log(1 + count)` | Right-skewed, with some zeros. |
| Dislikes | Use `log(1 + count)` | Right-skewed, with many zeros. |
| Favorites | Exclude | All observed values are zero. |
| Comments | Use `log(1 + count)` | Right-skewed, with zeros and a few large counts. |
I use `log1p()`, which calculates `log(1 + count)`. It reduces the skew
and works when the count is zero. None of the outcomes is used as is.
```{r}
outcomes <- setdiff(engagement, "favorite_count")
youtube[paste0("log_", outcomes)] <- lapply(youtube[outcomes], log1p)
```

## (c)

I fit a separate model for each outcome-characteristic pair, with `year`
included in every model. The binary predictors are coded as 0 and 1.
Rows with missing model variables are left out by `lm()`.
```{r}
features <- c("funny", "show_product_quickly", "patriotic",
              "celebrity", "danger", "animals", "use_sex")
youtube[features] <- lapply(youtube[features], as.integer)
results <- data.frame()
for (outcome in outcomes) {
  for (feature in features) {
    formula <- as.formula(paste0("log_", outcome, " ~ ", feature, " + year"))
    fit <- lm(formula, data = youtube)
    b <- summary(fit)$coefficients[feature, ]
    results <- rbind(results, data.frame(
      outcome = outcome, feature = feature,
      estimate = b[1], p_value = b[4], n = nobs(fit)
    ))
  }
}
knitr::kable(results, digits = 4, row.names = FALSE)
```
None of the seven characteristics is significant at 0.05 for any of the
four outcomes. There are therefore no significant positive or negative
associations to report.
For `log(1 + view_count)`, **danger** has the largest positive estimate:
**0.6668**, with **p = 0.0941**. The estimated association is positive after
adjusting for year, but it is not significant at 0.05.

## (d)

`danger` also has the smallest p-value for view counts, so I use it here.
The response is still `log(1 + view_count)`.
```{r}
view_results <- results[results$outcome == "view_count", ]
selected <- view_results$feature[which.min(view_results$p_value)]
selected
columns <- c("log_view_count", selected, "year")
d <- youtube[complete.cases(youtube[, columns]), columns]
X <- cbind(1, d[[selected]], d$year)
colnames(X) <- c("(Intercept)", selected, "year")
y <- d$log_view_count
beta_hat <- solve(t(X) %*% X) %*% t(X) %*% y
beta_hat
```
The three columns of `X` are the intercept, `danger`, and `year`.
The calculation uses $\widehat\beta=(X^{\mathsf T}X)^{-1}X^{\mathsf T}y$.
```{r}
fit <- lm(log_view_count ~ danger + year, data = d)
cbind(matrix = as.vector(beta_hat), lm = coef(fit))
all.equal(as.vector(beta_hat), unname(coef(fit)), tolerance = 1e-7)
```

